% arara: pdflatex: {files: [MathSACpr2014]}
% !arara: indent: {overwrite: yes}
\chapter{ALEKS pilot}\label{app:sec:aleks}
\section{MTH 20 Several classes during 2012--2013 AY (Edwards)}

The pilot includes the extensive use of ALEKS, a technology based assessment learning system,  in 2 on campus and 2 online classes each term.  

Course logistics: 
\begin{itemize}
	\item Students are walked through an introduction to the system and given an assessment
	\item Students are then provided with a very clear visual pie chart showing them what they know.
	\item ALEKS then provides students the opportunity to work on a range of instructor chosen topics at their current level.  
	Student only work on concepts they have not mastered.
	\item Explanations and videos are provided with each topic.
	\item Students are provided instant feedback and instant online teaching
	\item Students are not given the option to skip work that they have not mastered, essentially forcing them to learn the material and 
	fill in the concepts gaps that they began the class with.
	\item Students are routinely assessed with new topics available as they move through the course.
	\item Students are in the computer lab working on ALEKS throughout the class period
	\item Students, (generally for whom the material is recent) have the ability to move ahead.  
\end{itemize}

\subsection{Results and Statistics}
Reflecting only fall term math 20 students.  This was a definite pilot.  A variety of changes were incorporated into winter and spring terms.  Including additional lectures and assignments that had each class more closely resemble more traditional class.
\begin{itemize}
	\item Students loved the instant feedback
	\item Students enjoyed the ability to work in the ALEKS system, choosing their topics, and getting ahead when desired.  There were very, very few complaints about the system.
	\item Students became aware of how much time they studied, with a clear visual of the relationship between study time and learning. 
	\item FOUR students last term completed the math 20 material, moved on to math 60 material, took and passed my math 60 final exam.
\end{itemize}

On Campus Classes:  
\begin{itemize}
	\item 78\% of students passed math 20 last Fall compared to 89\% using ALEKS  (7am class result was 63\% passed using ALEKS)
	\item Of those that went on to math 60:  60\% passed last Fall compared to 69\% using ALEKS (7am class result: 13\% passed, 1 in 8)
\end{itemize}
DL Classes:
\begin{itemize}
	\item 62\% of students passed math 20 last Fall compared to 71\% using ALEKS
	\item Of those that went on to math 60:  61\% passed last Fall compared to 46\% using ALEKS.  
\end{itemize}

\section{Pilot in Math 112 during Winter 2013 (Louie)}
\fixthis{this paragraph is very long}
I think the most beneficial part about ALEKS is the instant feedback and instant teaching. This gives the student a chance to fill in the holes of their knowledge. My data was from a very small group. I compared one class (no aleks) to 2 classes (with classes). I averaged data from the 2 classes to get more accurate results. Surprisingly the grade distribution and overall pass rate was very close from ALEKS to no ALEKS. The distribution of grades was also very similar. In both classes my pass rate was 73\% which is well above the current 57\% campus average pass rate. The attrition rate for non ALEKS classes was 32\%.  The ALEKS class averaged a mere 14.7\%. Does ALEKS keep students on task and less likely to
withdrawal from the course? The numbers seems to support it but the sample size was small. The other benefit to ALEKS is requiring students to do homework and keeping track of their progress. The max average time spent on ALEKS was 15.4 hours and the minimum 1.6 hours per week. Despite my lack of data to support higher grades, I am confident that the students should be more prepared for Math 251. I am planning to check the success rates of the students who went on to calculus at Cascade campus. I would like to see if the pass rate of ALEKS students is higher than that of NON-ALEKS courses. Data is still in the works.

Best of all, students were forced to complete ALL homework and lectures seemed to flow with little interruption. I have not completely compiled the results from an ALEKS survey I gave at the end of the term. However the beginning results favor that most students enjoyed using ALEKS for homework and found it helpful in their learning.
