% arara: pdflatex: {files: [MathSACpr2014]}
% !arara: indent: {overwrite: yes}
\chapter[Faculty composition and qualifications]{Faculty:  reflect on the
composition, qualifications and development of the faculty}
%\epigraph{ I want to let you know that every single success in my life now is due in part to your teachings.  I can't thank you enough \& I hope that if for nothing else, you have made a great influence on me.}{Oke Tammik, PCC Mathematics Student, December 2013}

\epigraph{Professional development is particularly important for developmental
 education instructors, as these individuals tend to have limited previous
 training for teaching basic skills students. Unfortunately, studies have
 found that most community 
 colleges provide only episodic staff development activities, which tend to
 take the form of one-day workshops or seminars led by outside experts, or
 else isolated conversations among colleagues or departmental meetings that
 focus on logistics or content knowledge rather than pedagogy. Sadly, studies
 have revealed that such isolated professional development does little to
 change individuals' everyday practice, as they become ubsumed in normal
 routines and have little support for integrating new learning into their
 practice.}{(Rutschow, E.Z. \& Schneider, E. 2011. Unlocking the gate: What we
 know about improving developmental education. Oakland, CA: MDRC, p. 66)}

{\bfseries Provide information on each of the following}
\section{Quantity and quality of the faculty needed to meet the needs of the
program or discipline.}
The total number of full time faculty at all campuses between 2011 and 2013
varied between 36 to 41 and the 
part time faculty varied between 143 to 158 on any given term,  not including
Summer.     The percent of \emph{all} courses (pre-college and college level)
taught by full-time instructors during this time period varied from a low of
24.9\% at Cascade to a high of 41\% at Sylvania (see
\cref{reflect:tab:percentallcourses}).

From the academic year 2008/09 to 2012/13 there was a significant increase in the
number of students taking math courses at all campuses as shown
\cref{reflect:tab:enrollment}. 
\begin{table}[!htb]
  \begin{widepage}
	\begin{minipage}[t]{.4\textwidth}
		\centering
		\caption{Percentage of courses taught by full-time faculty from 2011--2013}
		\label{reflect:tab:percentallcourses}
		\begin{tabular}{rrrr}
			\toprule
			SY   & RC     & ELC  & CA     \\    
			41\% & 28.2\% & 26\% & 24.9\% \\
			\bottomrule
		\end{tabular}
	\end{minipage}%
	\begin{minipage}[t]{.6\textwidth}
		\centering
		\caption{Enrollment Difference from AY 08/09 to AY 12/13}
        \label{reflect:tab:enrollment}
		\begin{tabular}{lrr}
          \toprule
			Campus & Enrollment Difference & \% increase \\
            \midrule
			SY     & 5277                  & 48.59\%     \\
			CA     & 3666                  & 64.82\%     \\
			RC     & 5333                  & 55.87\%     \\
			ELC    & 3171                  & 103.09\%    \\
            \bottomrule
		\end{tabular}
	\end{minipage}
  \end{widepage}
\end{table}

\Cref{app:tab:analysisPTFT} summarizes the breakdown of courses taught by
full-time and part-time faculty from Summer 2011--Spring 2013; breakdowns 
by \emph{term} are given in \vref{app:sec:analysisPTFT}.

\begin{table}[!hb]
	\centering
	\caption{Summary of sections taught (by campus) from Summer 2011--Spring 2013}
	\label{app:tab:analysisPTFT}
	% summary
	% summary
	% summary
	\pgfplotstableread[col sep=comma]{./data/sectionsTaughtPTFT/sectionsTaughtPTFT2011-2013.csv}\sectionsTaughtSummary
	\pgfplotstabletypeset[sectionFTPT]{\sectionsTaughtSummary}
\end{table}

In reference to ``quality of the faculty needed to meet the needs of the discipline,''  it is insufficient to look at degree or experience qualifications alone.  Even a short list of what we expect from our mathematics faculty would include, but not be limited to that she/he:
\begin{itemize}
  \item possess an understanding of effective mathematics teaching methodologies and      strategies, and be able to adjust in response to student needs;
  \item  teach  the course content as outlined in CCOGs and with the appropriate mathematical  rigor;
\item show genuine commitment to students' success;
\item identify problems when students encounter difficulties learning;
\item demonstrate an ongoing intellectual curiosity about the relationship
  between teaching and learning;
\item manage classroom learning environments and effectively handle student
  discipline problems;
\item demonstrate technological literacy needed in the teaching of mathematics;
\item participate in professional organizations;
\item develop, evaluate and revise the mathematics curricula;
\item serve and contribute to the PCC community as a whole through campus and
  district wide committees and activities such as, for example, learning assessments committees or
  those that seek to reorginazate pre-college level math pathways.
\end{itemize}
In addition, with the enormous enrollment increases of the past several years,
there are more students than ever needing both remediation in mathematics and
guidance in general about what it takes to be a successful college student.

Addressing this section heading directly, the `quantity' of full-time faculty
needed to achieve the `quality' goals noted above is currently inadequate.  It
is primarily the full-time faculty that has the time, resources and
institutional support to fully realize the expectations noted above.  Part-time
faculty are dedicated, but the expectations are different given their
compensation.   To increase the probability that a student moves successfully
through our mathematics courses without sacrificing quality, having a larger
full-time faculty presence than currently exists is needed.
\fixthis{comment below}
%  maybe delete ``\ldotsgiven their compensation.'' or trade it for more words.
%  The true story is much greater than pay--it includes limited office space,
%  the necessity of travel between campuses or colleges, limited initial
%  knowledge of PCC support services, limited time or training to learn PCC
%  expectations of teaching practices, etc.   Personally I think you have an
%  opportunity here to fight for better support for your PT faculty as well.
%  At very least please don't reduce it to a matter of pay.

In recognizing the need for more full-time faculty, we do not want to downplay
the skills and talents of our part-time faculty.  We have approximately 150
part-time instructors that serve our students each term, many of whom have
teaching experience from other colleges and universities; they bring additional
experiences from industry, other sciences, high school and middle school
education, and so much more.  Since they teach such a high percentage of our
classes, their success is crucial to our students' success.
\fixthis{comment below}
% Here's your chance.  What do PT instructors need? Try ``To support PT faculty
% success at PCC, consider expanding NFI to include all new PT faculty, and
% changing the name of PT faculty institute to PCC Institute for Excellence in
% Teaching and invite ALL faculty to take part in the Saturday workshops each
% term.''  But these are my dreams.  Surely you have your own you could fill in
% there?

\subsection{Recommendation}
\recommendation{Given PCC's limited resources, the MTH SAC welcomes the input
of the administration to help us prioritize our various needs and find/develop
targeted and effective professional development for as many faculty members as
possible, including part-time faculty.  }
\fixthis{comment below}
% there is no recommendation here.  You could at least state the needs you
% somehow require administrative support to effectively prioritize on your
% behalf.  That or delete this non-section and save it for when you have
% something to share.

\subsection{Extent of faculty turnover and changes anticipated in the next five
years.} 
Since 2011, ten full-time instructors have been hired and eleven full-time
instructors have left campuses across the district (this includes full-time 
temporary positions).  Of the eleven full-time
instructors who left, five retired, one left to pursue other job opportunities,
and one returned to another teaching job after the term of the PCC position
ended.  Three of the retirements occurred at Sylvania and one each at Rock
Creek and Cascade.  Four full-time instructors left one campus in order to work
full-time at another campus.   There was one transfer from Cascade to Rock
Creek, one from Cascade to Sylvania, one from Rock Creek to Cascade, and one
from Rock Creek to Sylvania.  In the next five years, similar turnover with
full-time faculty is expected.
\fixthis{comment below}
% You haven't accounted for all 11 who left.  Maybe the rest is counted among
% the FT temporary faculty who returned to PT ranks?
% Doesn't this section ask for actual anticipated retirements among FT faculty,
% in addition to everything else you've discussed here?

Since 2011, 53 part-time instructors have been hired and 35 part-time
instructors have left campuses across the district.  Of the three campuses,
Rock Creek has the most part time faculty turnover, followed by Cascade and
Sylvania.  Reasons for leaving varied, but at least eight of the part-time
instructors who left campuses simply moved to another campus in the district
(see \vref{app:sec:facultyDegrees}).

\subsection{Extent of the reliance upon part-time faculty and how they compare
with full-time faculty in terms of educational and experiential backgrounds.}
Across the district, the mathematics departments rely heavily upon part-time
faculty to teach the majority of the math classes offered.  Between 2011 and
2013, 75.1\% of the classes at Cascade were taught by part-time instructors,
71.8\% at Rock Creek, 72.7\% at Southeast, and 59\% at Sylvania.  This reliance
on part-time faculty to teach classes has been a challenge to the departments
in a number of ways:  
\begin{itemize}
\item the turnover of part-time faculty is higher and
thus there is a need to orient new employees more frequently and provide
mentoring and guidance to them as well;
\item many part-time faculty are on
campus only to teach their courses, and thus often do not attend meetings and
keep up with current SAC discussions on curriculum.  
\end{itemize}
For these reasons, classes have a higher probability to be taught with less consistency than the
mathematics SAC would like.  Increasing the number of full-time faculty (and
thus decreasing the dependence on part-time faculty) would mitigate much of
this inconsistency; complete details are given in  \vref{app:sec:analysisPTFT}.

Part-time faculty educational backgrounds vary much more than the full-time
faculty backgrounds.  Full-time instructors have master's or doctorate degrees
in mathematics or related fields with extensive math graduate credits.  About a
quarter of the part-time instructors have bachelor's degrees and the rest have
either a master's or doctorate degree.  The part-time instructors come from a
variety of employment backgrounds and have different reasons for working
part-time.  They may be high school instructors (active or retired), may come
from a household in which only one member is working full time while the other
teaches part time, may be recently graduated MS or MAT students seeking full
time employment, may be working full time elsewhere in a non-educational field,
or may be retired from a non-educational field (see
\vref{app:sec:facultyDegrees}).
%  Is it true that a quarter of our PT faculty only have BS degrees?  How did
%  this happen?  I don't see room for that in the IQs.  How did you get this
%  data?  

\subsection{How the faculty composition reflects the diversity and cultural
competency goals of the institution.}
The mathematics SAC is deeply committed to fostering an inclusive campus
climate at each location that respects all individuals regardless of race,
color, religion, ethnicity, use of native language, national origin, sex,
marital status, height/weight ratio, disability, veteran status, age, or sexual
orientation.  Many of these human characteristics noted above are not
measurable nor necessarily discernible.  However, PCC does gather data on
gender and race/ethnicity, as detailed in \cref{reflect:tab:racialethnicmakeup}
(see also the extensive demographic data displayed in
\vref{app:sec:demographicdata}).

\begin{table}[!htb]
  \centering
  \caption{Racial/Ethnic Make-up of PCC Faculty and Students}
  \label{reflect:tab:racialethnicmakeup}
  \begin{tabular}{rrrr}
    \toprule
            &PT Faculty &   FT Faculty  & Students\\
            \midrule
    Male    & 54.1\%    & 53.2\%     & 55\% \\
    Female  & 45.9\%    & 47\%       & 45\% \\
    Asian /Pacific Islander & 7.7\% &   6.4\%    & 8\%\\
    Black or African American &   1.1\%  &  0.0\% &   6\%\\
    Hispanic/Latino &  2.2\%   & 4.3\%  &  11\%\\
    Multiracial & 1.1\%  &  0.0\%  &  3\%\\
    Native American & 0.0\%  &  0.0\%  &  1\%\\
    Unknown/International   & 12.2\% &  4.3\% &    3\%\\
    Caucasian   & 75.7\%   & 85.1\%   & 68\%\\
    \bottomrule
  \end{tabular}
\end{table}

Our SAC will continue to strive toward keeping our faculty body ethnically
diverse and culturally competent, but it is an area where improvement is
needed. In terms of hiring, there is a shortage of minorities in the Science, Technology, Engineering
and Mathematics (STEM) undergraduate and graduate programs, which makes our
recruitment of minority faculty difficult. \label{reflect:page:stem}

\subsection{Recommendations}
  \recommendation{Math chairs and deans should strongly recommend that full-time faculty
    attend workshops related to diversity and cultural competency issues.}

  \recommendation{Departments should be encouraged to provide diversity/cultural
  competency training for part-time faculty as part of their contractual
  meeting requirements.}

\recommendation{Hiring committees need to work with HR to identify
  and aggressively target mathematics graduate programs in the Northwest with
  minority students who are seeking teaching positions in community colleges.}

  \recommendation{Departments on all campuses should increase efforts to find candidates
    for the Faculty Diversity Internship Program \cite{affirmativeaction}.}

\section{Report any changes the SAC has made to instructor qualifications since
the last review and the reason for the changes.}
In Spring 2011, prompted by the transfer of Math 20 from Developmental
Education (DE) to the Mathematics SAC, the math instructor qualifications were
changed.  Math 20 had been the only remaining mathematics course in the DE SAC.  

The transfer included transitioning three full-time DE math instructors at
Sylvania into the Math Department at Sylvania.  At this time, instructor
qualifications for math faculty were examined and changed to reflect the
inclusion of DE math faculty.  It was determined that separate qualifications
should be written for pre-college and college level courses.  These
qualifications were written so that all of the full-time DE math faculty
transitioning into the math department (as well as any new DE math faculty
hired) were qualified to teach the pre-college level courses and any new math
faculty were qualified to teach all of the math courses.  

For instance, a masters degree in mathematics education (instead of just
mathematics) was included as an optional qualification for full-time
instructors teaching pre-college level courses.  Also a masters degree in
mathematics education became an option for part-time instructors teaching MTH
211--213 (the sequence for elementary education math teachers).  Additionally,
at the request of the administration, the terms `part-time' and `full-time'
were removed from instructor qualifications in order to satisfy accreditation
requirements.  Instead of labeling what had traditionally been part-time
qualifications as `part-time,' these qualifications were labeled `Criteria for
Provisional Instructors.'

In Winter 2013, the math instructor qualifications were again changed at the 
request of the math department chairs.  The `provisional' labeling
from the last revision had required math department chairs to regularly
re-certify part-time (`provisional') instructors.  In order to avoid this
unnecessary paperwork, the SAC adopted a three-tiered qualification structure
based on full-time, part-time, and provisionally-approved part-time instructors
(mainly graduate students currently working on graduate degrees).  The
part-time (non-provisional) tier was labeled `Demonstrated Competency.'
Complete details of instructor qualifications are given in \vref{app:sec:instructorquals}.
\fixthis{comment below}
% Also, it seems that related to your SY FT hires for next year we're going to
% have to address IQs again.  You may want to mention this here.

\section{How have professional development activities of the faculty contributed to the strength of the program/discipline? If such activities have resulted in instructional or curricular changes, please describe.}

The members of the mathematics SAC, full-time and part-time alike, are very
committed to professional development.  As with members of any academic
discipline, the faculty in the math SAC pursue professional development in a
variety of manners.  Traditionally these activities have been categorized in
ways such as `membership in professional organizations' or `presentations at
conferences'.  The members of the math SAC do not in any way devalue the
engagement in such organizations or activities, and in fact a summative list of
such things can be found in
\fixthis {insert appropriate appendix reference here}.

Nor do the members in any way diminish individual pursuit of professional
development.  In an attempt to acknowledge such pursuits, each member of the
full-time faculty was asked to submit one or two highlights of their
professional development activities over the past five years.  Those
submissions can be found in
\fixthis {insert appropriate appendix reference here}.

It should be noted that the list of organizations and activities found in these
appendices are not exhaustive; they merely connote a representative sample of
the types of professional development pursuits engaged in by members of the
math SAC.

The members of the math SAC realize that if there is going to be
institutional-level change that results in increased success and completion
rates for students enrolled in DE mathematics courses, there are going to have
to be targeted and on-going professional development activities with that goal
in mind and that all mathematics faculty, full-time and part-time, are going to
have to take advantage of those opportunities.  This is especially important
since the majority of our faculty do not have expertise in working with
developmental mathematics students. We look forward to working with the broader
PCC community as we pursue our common goal of increased student success and
completion, and we look forward to the college's support in providing
professional development opportunities that promote attainment of this goal.
\fixthis{would this be better as a recommendation? We look forward.... could be We recommend the college continue to support...}
