% arara: pdflatex
\documentclass{beamer}
\usepackage[latin1]{inputenc}
\usetheme{Warsaw}
\usepackage{standalone}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{mindmap}
\usetikzlibrary{decorations.text}

\tikzset{
    invisible/.style={opacity=0},
    visible on/.style={alt=#1{}{invisible}},
    alt/.code args={<#1>#2#3}{%
      \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
    },
  }
%Does someone have a theme they like or one they created for PCC?

\title{Mathematics SAC Program Review}
\institute{Portland Community College}
\date{January 31, 2014}
\begin{document}

\begin{frame}
\titlepage
\end{frame}

%Let me know if there is anything out of place or something you think should be moved to another page. 

\begin{frame}
	
	\begin{tikzpicture}[mindmap,
concept/.append style={fill={none}},
root concept/.style={concept color=blue},
level 1 concept/.append style=
{level distance = 35mm},
level 2 concept/.append style=
{level distance = 19mm},
every node/.append style={align=center,scale=0.7},
]
\node [concept,font=\huge] {Math SAC\\2008--2013}
child[grow=145, concept color=orange,visible on=<2->] {node[concept] {Action Committees} 
        child[grow=45, visible on=<2->]{node[concept] {Social Justice}}
        child[grow=135, visible on=<2->]{node[concept] {LAC, LAS}}
        child[grow=180, visible on=<2->]{node[concept] {DL Standing}}
        child[grow=225, visible on=<2->]{node[concept] {Dual Credit}}
            }
child[loosely dotted,concept color=blue,grow=215,visible on=<3->] {node[concept] {Faculty, Projects} 
        child[grow=135,visible on=<3->]{node[concept] {Study Skills}}
        child[grow=180,visible on=<3->]{node[concept] {WeBWorK}}
	child[grow=225,visible on=<3->]{node[concept] {Accessibility}}
	child[grow=270,visible on=<3->]{node[concept] {ALEKS, GeoGebra}}
	child[grow=315,visible on=<3->]{node[concept] {Involved Faculty}}
        }
child[dash pattern=on 1pt off 1pt on 1pt off 1pt, concept color=blue!50!purple,grow=35,visible on=<4->] {node[concept] {Curricular} 
	child[grow=0,visible on=<4-> ] {node[concept] {Created 111H}}
	child[grow=45,visible on=<4->] {node[concept] {Created 84}}
	child[grow=135,visible on=<4->] {node[concept] {Created 105}}
	child[grow=270,visible on=<4->] {node[concept] {Formed 111}}
	child[grow=315,visible on=<4->] {node[concept] {Adjusted 243}}
        }
child[loosely dashed,concept color=green!50!black,grow=325,visible on=<5->] {node[concept] {The Future} 
	child[grow=225,visible on=<4-> ] {node[concept] {Pathways}}
	child[grow=45,visible on=<4->] {node[concept] {AMP}}
	child[grow=-15,visible on=<4->] {node[concept] {ALC}}
	child[grow=-75,visible on=<4->] {node[concept] {WeBWorK}}
        };
%\node at (0,0) [inner sep=9mm,decorate,circle,decoration=
%{text along path,text={Equally Effective Equally Effective Equally Effective  Equally Effective }}] {};
\end{tikzpicture}

	
\end{frame}

\begin{frame}{Highlights from the last 5 years}

\begin{itemize}
%Is there a good way to reference the page numbers in the larger document? I thought it would be nice to have parenthetical page numbers. 

\pause \item Creation of Developmental Math Subcommittee
	\begin{itemize}
	\pause\item More meaningful outcomes
	\pause\item Research driven
	\pause\item Large faculty involvement
	\pause \item Large scale curriculum reform (future goal)
	\end{itemize}

\pause \item Study Skills Workbook
\pause \item Social Justice Workgroup
\pause \item Mathematics Accesibility Study
\pause \item WeBWorK
	\begin{itemize}
	\pause \item Accesibility
	\pause \item Free
	\pause \item Problem Library Development
	\pause \item Future Developments
	\end{itemize}

\pause \item Other Technology innovations in the classroom
	\begin{itemize}
	\pause \item ALEKS
	\pause \item GeoGebra
	\end{itemize}

\end{itemize}
\end{frame}

\begin{frame}{Highlights}
\begin{itemize}
\pause \item Formation of Math SAC Distance Learning Standing Subcommittee.
\pause \item Formation of Math Learning Assessment Subcommittee
	\begin{itemize}
	\pause \item Assessment Team
	\pause \item Action Team
\end{itemize}

\pause \item Faculty committed to professional development and serving the college. 
	\begin{itemize}
	\pause \item Multiple college wide committees (including several chairs)
	\pause \item Conference attendance
	\pause \item Memberships in professional organizations
	\pause \item Improved Dual Credit Relationships
	\end{itemize}

\end{itemize}
\end{frame}

\begin{frame}{Highlights}
\begin{itemize}

\pause \item Enrollment growth in LDC courses. 
\pause \item Curricular Changes
	\begin{itemize}
	\pause \item  Condensed MTH 111B and MTH 111C into MTH 111. 
	\pause \item MTH 105 replaced MTH111A and increased enrollment.  
	\pause \item MTH 111H created. 
	\pause \item All classes with MTH prefix are now part of the MTH SAC. 
	\pause \item MTH 243 credit change and future pre-requisite change. 
	\pause \item MTH 84 (LaTeX) course created
	\end{itemize}
\pause \item ALC and AMP
	\begin{itemize}
	\pause \item ALC (Alternative Learning Centers) courses allow students the opportunity to work on the math they need to work on for credit (but not for pre-requisites). 
	\pause \item AMP (Accelerated Math Placement) allows students to review material and gives them the chance to take the placement exam again for (possibly) higher placement. 
	\end{itemize}

\end{itemize}
\end{frame}
 
\begin{frame}{The Future of the Math SAC}

%insert flow chart here
%maybe this will work once it gets compiled?

\makebox[\textwidth][c]{%
  \input{graphics/pathwayDiagram.tex}
  }

*Every student in a given section of Math in Context I 
would also be enrolled in a common section of the CG 
course.  The instructors of a given pair of courses 
would work in a collaborative fashion and ideally visit 
one another's classes, especially the first week. 

Unlinked sections of the CG course would be offered 
for entry level DE mathematics students whose initial 
placement is `above' Math in Context I.  Students who 
pass Math in Context I but not the attendant CG course 
might be required to retake the course.

\end{frame}

\end{document}
