% arara: pdflatex: {files: [MathSACpr2014]}
\chapter[Outcomes and Assessment]{Reflect on learning outcomes and assessment, teaching methodologies, and content in order to improve the quality of teaching, learning and student success.}

\section[Course-Level Outcomes]{Identify and give examples of assessment-driven changes made to improve attainment of course-level student learning outcomes.   Where key sequences exist, also include information about assessment-driven changes to those sequences.}

There has been little direct assessment of course-level outcomes in the last five years for the following two reasons: 
\begin{enumerate}
\item The Curriculum Committee currently requires an ``out there'' focus for course-level learning outcomes, with no requirement that outcomes are assessable or measurable.\fixthis{Have asked Michele for a clarification of what "out there" means}
\item The annual assessment reporting for the Learning Assessment Council (LAC) has focused on the college's core outcomes, not course outcomes.  
\end{enumerate}


However, we are fortunate to have math faculty involved with the Learning Assessment Council (LAC) and the Curriculum Committee. This involvement has kept our SAC aware of the college's ongoing discussions regarding a possible future change for the focus of course-level learning outcomes  (i.e., expectation of measurability) and related accreditation standards (e.g., \cite[Standard 4.A.3]{NWCCU}).

Many of our current learning outcomes were developed to satisfy the requirements of an ``out there'' focus and this has resulted in oddly worded or aspirational outcomes.  Here are two examples from MTH 251 (Calculus 1):

\begin{quote}
Appreciate derivatives and limit-related concepts that are encountered in the real world, [and] understand and be able to communicate the underlying mathematics involved to help another person gain insight into the situation.
\end{quote}
\begin{quote}
Enjoy a life enriched by exposure to Calculus.
\end{quote}

While we hope that our students will be able to ``help another person gain insight into the situation'' and that they will ``enjoy a life enriched by exposure to Calculus,'' we recognize that outcomes like these are not easily measured. Before our SAC can make assessment-driven changes to improve students' attainment of outcomes, we need to first develop measurable outcomes that represent the intent of our courses.

In 2012/13 the MTH 60/65 CCOG subcommittee decided to develop course-level outcomes that were meaningful and assessable; this was done despite concerns that Curriculum Committee might reject them because they lacked the required ``out there'' focus.  The Math SAC was largely supportive of the shift to assessable outcomes, suggesting that we are supportive of the college transitioning toward a culture assessable outcomes at all levels.  

While crafting the new (\emph{draft}) course outcomes, the MTH 60/65 CCOG subcommittee connected each proposed outcome to an actual assessment activity used by a member of the committee in order to ensure that each outcome was truly measurable.  What follows are a few of the resultant (\emph{draft}) outcomes.
\begin{description}
\item[Argument Construction:] Construct and judge the validity of an argument. (e.g., Why does a particular symbolic representation match a particular graphical representation?)

\item[Representational Fluency:] Demonstrate the ability to distinguish different meanings of `variable', e.g., a variable can represent a varying quantity in an expression or an unknown quantity in an equation.

\item[Problem Solving:] Use appropriate (mathematical) tools in the context of problem solving, modeling, interpreting, etc. (Know what approach to take, what information you have, what information you need, what techniques you have to solve the problem, what the graph tells you, what the formula tells you, what model you can build).
\end{description}

The current layout of the CCOGs does not differentiate critical content from less critical content.  To address this, the committee discussed the option of formatting the content area of the CCOGs in a pyramid structure with the most critical content highlighted at the bottom of the pyramid.  

The current CCOGs also do not align the content to the course outcomes. To help make the CCOG a better communication tool, the committee discussed making explicit connections between the course content and the course outcomes as well as explicit connections between the course outcomes and the college's core outcomes.  This work was in progress when we realized that this shift in outcomes should not be done in isolation from other courses in the pre-college math sequence; in response, we formed a DE Math Subcommittee with the goal of creating vertically aligned outcomes that created a coherent progression from MTH 20 through MTH 95.

In summary, we are interested in developing rich, meaningful, and measurable outcomes that better represent the intent of our courses than our current outcomes do. Since course outcomes are required in syllabi, properly representing the intended focus for the course in the the course outcomes is critical. We hope that the conversations in the LAC and Curriculum Committee concerning course-level learning outcomes will lead to Curriculum accepting more flexible wording of course outcomes in the near future.\fixthis{Is there a Recommendation here?}

\section{Addressing College Core Outcomes}

\subsection{Describe how each of the College Core Outcomes are addressed in courses, and/or aligned with program and/or course outcomes.}

The colleges Core outcomes may be found at \cite{coreoutcomes}.

\begin{description}
\item[Communication] is stated in many of our CCOGs as a course outcome. We believe it is important for students to communicate ideas using mathematics in a meaningful manner through appropriate use of notation and concise accurate statements.  Here are some excerpts from course outcomes in MTH CCOGs: 
\begin{aquote}{MTH 111-112, MTH 211-213, MTH 251-254}
{\ldots}and then interpret and clearly communicate the results. 
\end{aquote}

\begin{aquote}{MTH 243-244}
{\ldots}and clearly interpret the results via written or oral communication. 
\end{aquote}

\begin{aquote}{MTH 111-112, MTH 243-244, MTH 251-254}
{\ldots}understand and be able to communicate the underlying mathematics involved to help another person gain insight into the situation.
\end{aquote}

\item[Community and Environmental Responsibility] is not directly addressed in our courses or outcomes. Our Social Justice Workgroup (\vref{cur:sub:socialJustic} and \vref{app:sec:socialJustic}) discusses issues related to teaching with community and environmental responsibility in mind. Also, sections of MTH 111H (\vref{cur:sub:111H}) have a related component involving tutoring math to someone in a student's community. \fixthis{it doesn't know how to refer to subsections}

\item[Critical Thinking and Problem Solving] is fundamental to mathematics, so at first glance it may seem that mathematics would be an easy fit for this core outcome.  However, traditionally most pre-college and lower division college courses treat math as though it is a mostly a procedural activity and there is very little focus on critical thinking and problem solving.  Recent discussions concerning our assessment of this core outcome have broadened our view and helped motivate a deeper consideration about what Critical Thinking and Problem Solving should entail for mathematics.  Some ideas that are being considered are:
\begin{itemize}
\item inquiry-based or project-based learning
\item problems with too much or missing information that has to be identified (just as the problems we encounter in the real world)
\item pattern recognition
\end{itemize}

\item[Cultural Awareness] is not formally covered in any of our courses.  Some individual instructors may incorporate elements of cultural awareness into their classes.  Some examples of how cultural awareness could be addressed by individual math instructors instructors are: 
\begin{itemize}
\item differences in notations and procedures used by different cultures
\item similarities and differences with regard to the use of mathematics as a tool 
\item attitudes toward mathematics in various cultures (e.g., math anxiety is not a worldwide issue) 
\item history of mathematical  development in various cultures and how it was influenced by dominant philosophical attitudes of the region
\item use of mathematics as an agent of social change by shining a light on inequity with regard to treatment of, and resource allocation for, marginalized populations as compared to the dominant culture (the social justice workgroup considers this) (see \vref{cur:sub:socialJustic} and \vref{app:sec:socialJustic}.) 
\end{itemize}

\item[Professional Competence] is described by PCC as the ability to ``demonstrate and apply the knowledge, skills and attitudes necessary to enter and succeed in a defined profession or advanced academic program.'' To obtain a four-year degree from a public university in the State of Oregon, almost every major requires that students complete at least one college-level math course.  As such, successful completion of a math course is an indicator of professional competence.  Similarly, many CTE programs have a MTH requirement for their program, so successful completion of a math course is also an indicator of those students' professional competence. 

We offer one sequence that directly addresses professional competence for one group of students.  Our  MTH 211-213 (Foundations of Elementary Mathematics) sequence of courses are taken by students interested in pursuing a career in teaching in the K-12 education system.  At minimum, two of the three courses are prerequisites for obtaining a degree in Education in Oregon.  Each course emphasizes specific topics of mathematical theory that are the basic building blocks of mathematics instruction in the K-12 system.  Additionally, students in some classes are required to maintain a portfolio, learn multiple assessment techniques, do field observations of teachers in the K-12 system, learn about the Common Core Standards and trends in education both state and nation-wide, and are given guidance in areas such as preparing for the CBEST and PRAXIS as well as decision making in the various avenues for pursuing a degree in education.  This exposure to what the field of education actually entails  helps students make sound career choices early in their academic pursuits.

\item[Self-Reflection] is an outcome that needs more discussion within the SAC to develop a shared understanding of what it means in the context of a mathematics course.  Many SAC members believe that helping students develop strong study skills will involve improving their ability to self-reflect; consequently, the SAC has discussed adding study skills as formal components of MTH 20 (see \vref{cur:sub:mth20}).  Also,  see \vref{cur:sub:studyskills} for more information on study skills material that incorporates student self-reflection. 

In 2011/12 the Math Learning Assessment Subcommittee (Math LAS) explored Self-Regulated Learning where students were guided in a self-reflective process that helped them evaluate their depth of understanding (or lack thereof).  Although Self-Regulated Learning was not incorporated into the assessment activity (due to the complexity and the lack of time remaining), members of the members of Math LAS who explored this believe it is worthy of future consideration.

\end{description}

\subsection{Update the Core Outcomes Mapping Matrix for your SAC as appropriate.}\fixthis{this link was part of the subsection but its inclusion screws up the toc file. Not sure what to do.}\url{http://www.pcc.edu/resources/academic/core-outcomes/mapping-index.html} 


\fixthis{link to core outcome mapping matrix in appendix}
\fixthis{It appears this section is blank? Have asked Michele for clarification.}


\section[Assessment of College Core Outcomes]{For Lower Division Collegiate (Transfer) and Developmental Education Disciplines:  Assessment of College Core Outcomes    (note:  Please include the full text of your annual reports as appendices, and summarize them here).  }\fixthis{Do they literally want these as appendices, or are the links good enough?}

Our full annual Learning Assessment reports can be found at these links. 

\begin{description}
\item[2009/10:] \url{http://www.pcc.edu/resources/academic/documents/math.pdf}
\item[2010/11:] \url{http://www.pcc.edu/resources/academic/documents/LDCMTHCritThinkCommunicReport2011.pdf}
\item[2011/12:] \url{http://www.pcc.edu/resources/academic/documents/Annual_Report_on_Assessment_MATH_2012_SR_PC.pdf}
\item[2012/13:] \url{http://www.pcc.edu/resources/academic/documents/MTH_2013_report_math.pdf}
\end{description}

\subsection{Describe the assessment design and processes that are used to determine how well students are meeting the College Core Outcomes}

\begin{description}
\item [2009/10: Critical Thinking  \& Problem Solving]

This was the first year the LAC asked SACs to assess a core outcome. Our assessment activity was developed by a small group of interested math faculty with minimal coordination with the full SAC. The student work that was obtained came from student volunteers enrolled in sections with the involved faculty; it was not a statistically sound sample.  The activity involved both direct and indirect assessment: 
\begin{itemize}
\item The direct assessment involved finding and correcting conceptual, arithmetic and formatting mistakes in expected procedural skills for MTH 65 and MTH 95.  The ``find and correct''\fixthis{someone who was involved with this should make this sentence make sense; have asked Michele} is requires the students to know how to do the problem and to evaluate what was incorrect.  This is a higher-order skill on Bloom's Taxonomy. 

\item The indirect assessment involved asking students to respond to questions like, ``Do you feel this class has improved your critical thinking and problem solving skills?''
\end{itemize}

\item[2010/11:  Critical Thinking \& Problem Solving and Communication]

At the beginning of 2010, the Math SAC decided to create a standing committee to ensure assessment work was a high priority. The Math Learning Assessment Subcommittee (Math LAS) was born.  There were 14 members, most of whom were full-time faculty. This was a big improvement from last year's work where a small group of faculty participated.

Although our previous year's Critical Thinking \& Problem Solving activity had been high on Bloom's Taxonomy, we decided that procedural skills did not fully capture how we want our student to think critically about mathematics.  Instead of continuing with the previous year's work, we decided to develop a new activity for Critical Thinking \& Problem Solving. 

Still very new to assessment, we did not know what type of activity might generate the most useful information.  To help us decide, we developed three assessment activities and collected student artifacts for each activity.  The chosen activity was randomly given to 12 of the 72 sections of MTH 65 held in Winter 2011.  (Note: We also created an activity for MTH 244, but there was an error in one of the questions.  Although a portion of the artifacts were evaluated, we ultimately decided to abandon this attempt and focus our limited resources toward the MTH 65 analysis.)

For the MTH 65 assessment, we collected 240 student artifacts.  To help ensure that data would be a true SAC-level assessment (vs.\ an evaluation of individual instructors), faculty members were instructed to remove identifying information from student work and submit their artifacts to an administrative assistant who tracked submission of work only.  Sixteen members of the SAC were normed to the rubric that had been developed by the Math LAS.  Two members of the LAC's Program Assessment for Learning (PAL) facilitated this work and guided faculty through a trend analysis.  (Note: Our process and involvement of Math SAC members was so impressive as compared to other large SACs that the Learning Assessment Council awarded the Math SAC an ``Oscar'' at their Spring circus event.)

\item[2011/12: Self Reflection and Professional Competence]

This year we sent a survey to all students enrolled in a math class in the first week of Spring 2012.  The Math LAS was awarded a Learning Assessment Council grant and we used these funds to hire a consultant, Una Chi, to help us refine the survey and evaluate the student responses to the survey.  We also discussed the wording of the questions with the DE Reading and Writing faculty members to help ensure the questions would be understood by all students. Approximately 2300 students respond to the survey, and the response rates for particular courses mirrored student enrollment in those courses and other demographic information.  The survey was an indirect measure of students' perceptions.  

For Self Reflection, we focused on questions that we felt would fit the following three areas:
\begin{enumerate}
\item Reflection -- Core reflective thinking items; autonomy and relatedness aspects from self-determination theory (see \cite{selfdetermination})
\item Orientation -- Mastery/performance, internal/external locus of control (hold self responsible vs. holding others responsible)
\item Competency -- Belief about self-ability to perform in math (3rd aspect from Self Determination Theory) \fixthis{REF?}
\end{enumerate}
Sample Self Reflection items on the survey:
\begin{itemize}
\item I know when I need help on a math concept.
\item When I get a math test back, my grade is what I expect it to be.
\item My feelings about math affect my learning of math.
\end{itemize}

For Professional Competence, we used the suggestions of our LAC coach to craft questions about students' perceptions of math in terms of their future job/career goals. 

Sample Professional Competence items on the survey:
\begin{itemize}
\item The skills I learn in a math class are not important to me or my future goals -- I just need to pass the course.
\item In PCC math classes, what knowledge, skills, habits or ways of thinking have you practiced that might help you in the work place? [Choices included punctuality, problem solving, working in groups, self discipline, career specific math skills, interpret graphs/charts]
\item My career interest requires some mathematical knowledge.
\end{itemize}

\item[2012/13:  Critical Thinking \& Problem-Solving and Professional Competence]

This was our third attempt on Critical Thinking \& Problem Solving.  Although our previous attempt lead to rich data, this year we decided to investigate students' ability to solve nine math problems that represent the topics covered in MTH 95 that we consider essential for success in MTH 111.  The assessment activity was administered to every face-to-face section of MTH 95 at all PCC campuses in the Winter of 2013.  We collected 677 student responses from 33 different sections across the college; all identifying information for both instructors and students were removed.

The math problems included in the assessment were selected by faculty from our current MTH 95 textbook.
\begin{itemize}
\item For Critical Thinking \& Problem Solving: We incorporated problems that contain units and involve a real-world context.
\item For Professional Competence:  We incorporated problems that emphasize the content needed to be successful in the next course, MTH 111.  For this activity, Professional Competence was interpreted as the, ``knowledge, skills and attitudes necessary to enter and succeed in a defined profession or advanced academic program'' (this quote is from description of Professional Competence found at \url{http://www.pcc.edu/resources/academic/core-outcomes/index.html}).
\end{itemize}
During the LAC's summer peer review process, our report won awards in two categories: ``Assessment Design'' and ``Planned Improvements to Increase Student Attainment of Outcomes''.  The awards were announced at 2013 SAC Chair Inservice; here is a link to the list of awards: \url{http://www.pcc.edu/resources/academic/learning-assessment/sac-resources.html#assessmentwin}.

\end{description}

\subsection{Summarize the results of assessments of the Core Outcomes}

Below we summarize the results of our annual assessment projects in academic years 2009/10, 2010/11, 2011/12, and 2012/13.

\begin{description}
\item[2009/10: Critical Thinking \& Problem Solving]
We did not evaluate the student artifacts due lack of time during the 2009/10 academic year. We intended on completing the work during 2010/11; however, since we did not have a statistically significant sample (only a few sections of student artifacts were collected and those sections were taught by the faculty members who were involved in the work rather than a randomized sampling) and since we wished to try a different type of assessment during 2010/11, we decided to not evaluate these artifacts.  Even though we did not evaluate student learning, the faculty members gleaned valuable information about the process of assessment. The assessment artifacts have been saved in case the Math LAS wishes to review them to guide future work.
\item[2010/11:  Critical Thinking \& Problem Solving and Communication]
\Cref{ass:tab:201011scores} shows the results of the rubric scores for our 2011 assessment of Critical Thinking \& Problem Solving and Communication in 13 sections of MTH 65.
\begin{table}[!htb]
\centering
\caption{2010/11 Assessment Scores}\label{ass:tab:201011scores}
\begin{tabular}{llll}
\toprule
Rubric Score & 1 or 2 & 3 & 4 or 5\\
&(below expectations)&(met expectation)&(exceeded expectation)\\
\midrule
CT\&PS &55\%&35\%&10\%\\
Communication &50\%&28\%&22\%\\
\bottomrule
\end{tabular}
\end{table}

Below is a brief analysis of the results (see report for more information; a link to the report is above, directly under 2C):\fixthis{ref}

We realized that the rubric scores did not tell us specific information (e.g., why did the artifact score ``below expectation'' or ``exceeded expectation''?).  The LAC Program Assessment for Learning (PAL) facilitators suggested that we do a trend analysis which produced more meaningful information. Here are some results of the trend analysis:
\begin{itemize}
\item Many students did not seem to realize that not all data are linear.
\item Many students were not able to give a well-supported conclusion.
\item Students typically do not represent equivalence correctly.
\item Many students incorrectly applied the idea of percentage.
\end{itemize}

\item[2011/12: Self Reflection and Professional Competence]
The results of the survey are too large to summarize.  Please see \#3 on the report.  (see link to the report above, directly under 2C)\fixthis{ref} 

Below is a brief analysis of the results (see report for more information):

For Self Reflection: 
\begin{itemize}
\item Students are not self-critical enough
\item There was a significant group mean difference between self-reported grade and self-reflection behavior on all grade level differences.  Note: ``grade level'' is the self-reported grade (A, B, C, D, F, P, NP, other) for the student's previous math class.
\item There is a clear difference in the reflective thinking ability of students in high-level math courses versus students in low-level math courses (like MTH 20).
\end{itemize}
For Professional Competence:
\begin{itemize}
\item Surprising that Engineering was chosen by 11\% of respondents. Nursing and Business were both at 7\%.\
\item Presentation skills ranked low by students. This may be ok, but would be a good SAC discussion of how important this is for our students in our classes. We may not wish to ``force'' students to present in lower level courses where math anxiety is increased.
\end{itemize}

\item[2012/13:  Critical Thinking \& Problem-Solving and Professional Competence]

Our assessment consisted of nine math problems.  After all submissions were graded, the average score was approximately 3.8 out of 9.  On a class-to-class basis the average had a low of 1.7 out of 9 to a high of 5.4 out of 9.  We were unsure of why the average of all classes was so low; nonetheless we found it alarming.  Did students not take the activity seriously since most instructors did not assign points to it?  Would it have been better to give it with the final exam when most students were prepared for a mathematics assessment? 

Data is shown more clearly in the report under \#3 in table and graphical format (see the link to the report above, directly under 2C)\fixthis{ref}

Below is a brief analysis of the results (see report for more information).
\begin{itemize}
\item The problems that involve working with function notation were answered correctly at a much lower rate than we expected given the amount of time dedicated to that topic in MTH 95.
\item A problem that involved linear equations was also answered correctly at a much lower rate than we expected as that topic is covered in MTH 60, 65 and 95.
\item Given that it was difficult to discern a student's conceptual understanding of a topic due to time constraints for the assessment and our decision to mark answers as correct or incorrect (with no ``partial credit''), we should consider altering this activity if used again.
\end{itemize}


\end{description}

\subsection{Identify and give examples of assessment-driven changes that have been made to improve students' attainment of the Core Outcomes.}

Below we discuss the assessment-driven changes that came out of our annual assessment projects in academic years 2009/10, 2010/11, 2011/12, and 2012/13:

\begin{description}
\item[2009/10: Critical Thinking \& Problem Solving]

We chose to start anew rather than analyze data that was not statistically significant.  No SAC assessment-driven changes from this year's work.

\item[2010/11:  Critical Thinking \& Problem Solving and Communication]

During 2010/11, it became clear that the Math LAS members would not be able to develop/conduct assessment and lead the SAC with assessment-driven changes from the prior year's work.  At the end of this academic year, the SAC created another assessment standing committee, called the Action Subcommittee.  This subcommittee will take the results and the recommendations from the prior year's work by the Math LAS and lead the SAC in deciding what should be changed and implementing the change.  For 2010/11 work, only individual changes to instruction occurred from the assessment results.  (More information on this can be found in question \#1 from our 2012 assessment report (see the link to the report above, directly under 2C) : \fixthis{ref}

\item[2011/12: Self Reflection and Professional Competence]

The Action Subcommittee brainstormed a list of actionable items from the 2012 research and decided work on the following item: ``disseminate successful ideas already used by our faculty for improving self-reflection via study skills and student-centered learning.''  The goal was to create activities that faculty could easily incorporate into their classes that would help students develop self-reflection behaviors that would lead to better study skills.  Math SAC members were asked to submit activities that were already being used successfully, and the committee received 25 different activities.  During an all-day SAC meeting we split into breakout sessions and each group was asked to look over the activities and create a list of best practices for incorporating them into the classroom. These worksheets are available online for instructors to download an incorporate into their classes (\url{https://sites.google.com/a/pcc.edu/self-centered-learning-study-skills-for-mathematics}).  Additionally, \fixthis{ref}< REF: Jessica's study skills as mentioned in a different part of this Program Review: see 3E>, a faculty member created a series of self-reflection and study skills videos that are being used in a lot of developmental classes.

\item[2012/13:  Critical Thinking \& Problem-Solving and Professional Competence]

Below are some of the suggested actionable items suggested by the Math LAS from the 2012/13 assessment work  (For the complete list see \#4 in our 2013 assessment report: <(see the link to the report above, directly under 2C) : \fixthis{ref}
\begin{itemize}
\item Add to the CCOGs the expectation that students check the reasonableness of their results (e.g., a result of $-5$ or $1{,}000{,}496$ would not be a reasonable result for ``the number of hours driven on a weekend trip'').  Ideally, students should be encouraged to develop a habit of verifying their results regardless of whether or not the problem is a contextualized problem or not.
\item Create a minimum skills test for MTH 95.
\item Discuss methods of course content delivery in a way that supports both full- and part-time faculty.
\item Form a Developmental Math Committee that will research different ways we might be able to redesign our pre-college curriculum in order to better prepare students for college-level math as well as better serve students in CTE programs.
\end{itemize}

\end{description}

The Action Subcommittee is currently reviewing the 2012/13 assessment work and may propose to the SAC other ideas for implementation.


