% arara: pdflatex: {files: [MathSACpr2014]}
% !arara: indent: {overwrite: yes}
\chapter{Changes in ALC courses}\label{app:sec:alc}
After MTH 20 was moved to the Math SAC, the ALC Math courses were the only math courses left in the DE SAC. The ALC instructors therefore requested that these courses too would be moved to the Math SAC. After the DE and Math SACs gave their support, the courses were moved in January 2013.

Historically, the ALC math courses have only included curriculum up to MTH 65, but after the move was completed, the Math SAC voted to also include MTH 95 curriculum from Fall 2013 on.

Furthermore, the ALC math courses have been impacted by the new no-repeat policy. Historically, these courses could be repeated many times because they included three levels (now four). Since Winter 2014, each level has their own set of courses. Following is a listing  of the previous and the new/changed courses:

Previous courses (until Fall 2013):

ALC 60 ``Basic Math Skills Lab''

ALC 61 ``Basic Math Skills Lab''

ALC 62 ``Basic Math Skills Lab''

ALC 63 ``Basic Math Skills Lab''

New courses (since Winter 2014):
\fixthis{formatted differently? table? not sure}

ALC 20A ``Math 20 Lab - 0 credits'' \\
ALC 20B ``Math 20 Lab - 1 credits'' \\
ALC 20C ``Math 20 Lab - 2 credits'' \\
ALC 20D ``Math 20 Lab - 3 credits'' \\

ALC 60A ``Math 60 Lab - 0 credits'' \\
ALC 60B ``Math 60 Lab - 1 credits'' \\
ALC 60C ``Math 60 Lab - 2 credits'' \\
ALC 60D ``Math 60 Lab - 3 credits'' \\

ALC 65A ``Math 65 Lab - 0 credits'' \\
ALC 65B ``Math 65 Lab - 1 credits'' \\
ALC 65C ``Math 65 Lab - 2 credits'' \\
ALC 65D ``Math 65 Lab - 3 credits'' \\

ALC 95A ``Math 95 Lab - 0 credits'' \\
ALC 95B ``Math 95 Lab - 1 credits'' \\
ALC 95C ``Math 95 Lab - 2 credits'' \\
ALC 95D ``Math 95 Lab - 3 credits''
