% arara: pdflatex: {files: [MathSACpr2014]}
% !arara: indent: {overwrite: yes}

\chapter{Core Outcomes Mapping}\label{sec:app:coreoutcomes}
\begin{widepage}
\begin{minipage}[t]{.5\textwidth}
Mapping Level Indicators:
\begin{enumerate}[start=0]
  \item Not Applicable.
  \item Limited demonstration or application of knowledge and skills.
  \item Basic demonstration and application of knowledge and skills.
  \item Demonstrates comprehension and is able to apply essential knowledge and skills.
  \item Demonstrates thorough, effective and/or sophisticated application of knowledge and skills.
\end{enumerate}
\end{minipage}\hfill
\begin{minipage}[t]{.34\textwidth}
Core Outcomes: 
\begin{enumerate}[label=CO\arabic*]
\item Communication.
\item Community and Environmental Responsibility.
\item Critical Thinking and Problem Solving.
\item Cultural Awareness.
\item Professional Competence.
\item Self-Reflection.
\end{enumerate}
\end{minipage}%
\end{widepage}

\vfill
MTH courses 10B, 10C, 11B, 11C, 15, 20B, 21C, 22, 22C, 23C, 24C, 27C, 75, 91, and 92 remain in the catalog, but are no longer offered and have not been offered since at least Spring 2012.
\pgfplotstableread[col sep=comma]{./data/outcomeMatrix.csv}\coreOutComes

\footnotesize
\pgfplotstabletypeset[
begin table=\begin{longtable},
end table=\end{longtable},
	every head row/.style={
		before row=\toprule,
		after row=\midrule},
	every last row/.style={after row=\bottomrule},
    every row no 3/.style={after row=\cmidrule{2-8}},
    every row no 12/.style={after row=\cmidrule{2-8}},
    every row no 15/.style={after row=\cmidrule{2-8}},
    every row no 21/.style={after row=\cmidrule{2-8}},
	columns/course/.style={string type,column type=l},
	columns/name/.style={string type,column type=l},
]{\coreOutComes}
\normalsize